### 2024-06-03 manuscript2.Rmd
3/6

---
title: "Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study"
output: "html_document"
author: "Elin Sun Cao"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study
### Background
Trauma is a leading cause of morbidity and mortality worldwide. A cornerstone in trauma quality improvement programs is multidisciplinary morbidity and mortality review The purpose of these reviews is to identify opportunities for improvement (opportunities for improvement) and corresponding corrective actions. Many patients with severe trauma are admitted to the intensive care unit (ICU), but little is known about opportunities for improvement in this cohort. The aim of this study was to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

### Methods
A registry-based study of trauma patients admitted to the ICU between 2014-2023 at the Karolinska University Hospital in Solna, Sweden, who were reviewed regarding the presence of opportunities for improvement in a morbidity and mortality review. Univariate and multivariable logistic regression was used to determine associations between selected patient and process factors and opportunities for improvement.

### Results
opportunities for improvement was identified in 143 (9,9%) out of 1449 patients. Mechanical ventilation > 7 days (OR 0.35; 95% CI 0.10, 0.89; p=0.05), Revised Trauma Score (OR 1.21; 95% CI 1.08, 1.37; p=0,002), > 7 days in the ICU (OR 1.93; 95% CI 1.23, 3.03; p=0.004) and survival 30 days after injury (OR 3.29; 95% CI 1.60, 7.54; p=0.002) were significantly associated with increased odds of opportunities for improvement.

### Conclusion
Several patient and process factors were found to be associated with opportunities for improvement, indicating that patients who were relatively more stable upon arrival, patients with prolonged ICU stay, patients with prolonged mechanical ventilation, and patients who were alive 30 days after injury had higher odds of opportunities for improvement.

# Introduction

Trauma — the clinical entity of external injury and the body’s associated response — is a leading cause of morbidity and mortality worldwide [@60, @6]. Patients with major trauma often require intensive care unit (ICU) admission, where managing these patients can be challenging and result in complications or death [@64, @69]. While these complications are often expected, the complexity of ICU-care requires providers to be highly adaptable and makes it prone to mistakes and errors. Peer review is a cornerstone of trauma quality improvement programs to identify preventable causes of morbidity and mortality in trauma patient [@70], referred to as opportunities for improvement [*ref* - 10 i texten]. Previous research on opportunities for improvement in trauma care has often focused on the initial management and have identified airway and ventilator management, fluid resuscitation, delays to treatment and judgement errors as common areas for opportunities for improvement [*ref 10*].

Prior research has focused on classifying errors in preventable trauma deaths, with clinical performance errors, often stemming from a lack of knowledge, being identified as the most common type, particularly in the emergency department [@62]. However, reporting of errors in trauma resuscitation varies between settings and studies [@63]. As a results, there is a notable gap in research regarding error reporting practices in trauma care — especially for trauma patients admitted to the ICU [@12, @64]. Assessing preventable deaths among ICU-admitted patients poses a significant challenge due to their medical complexity, making it difficult to establish the standard of care for each individual. Moreover, despite receiving optimal treatment, these patients remain at elevated risk for complications and increased mortality rates [@9]. While numerous studies have explored various facets of trauma care enhancement, research specifically centered on opportunities for improvement in ICU-admitted trauma patients remains limited [@64]. The aim of this study was to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

# Methods

## Study design

We conducted a retrospective single-center cohort study based on data from the Karolinska University Hospital trauma registry, which is included in the Swedish Trauma registry (SweTrau), and the hospital's local trauma care quality database.

## Setting

The Karolinska University Hospital in Solna is the trauma center designated to receive all severely injured patients in the wider metropolitan area of Stockholm [@17]. The hospital has direct access to radiology, surgery, intensive care, and consultants from relevant specialities [@18]. The care of adult trauma patients is screened for opportunities for improvement using a combination of audit filters and individual review by specialized nurses. Those identified as having a higher likelihood of opportunities for improvement are discussed in a multidisciplinary peer-review held every 6-8 weeks. The presence or absence of opportunities for improvement is decided among the participants of the review and subsequently documented in the trauma care quality database. Before 2017 a non-systematic selection from clinicians involved in the opportunities for improvement process was done by filtering for patients that were thought to have a high risk for opportunities for improvement, but starting in 2017 the care of all patients is screened.

## Participants

The trauma registry includes all patients admitted with trauma team activation, regardless of Injury Severity Score (ISS), as well as patients admitted without trauma team activation but found to have an ISS of more than 9. We included all patients who had been included in the opportunities for improvement screening process between January 1, 2014, and February 1, 2023, and who were also admitted to the ICU. We excluded patients who were younger than 15 years and patients who were dead on arrival. Patients under 15 years of age were not included in the study as their clinical and review pathway differ compared to adults.

## Study size

When conducting logistic regression analysis, the number of events per variable (EPV) equal to 10 is a rule of thumb. If the EPV falls below this threshold, the results for regression analyses should be interpreted with caution [@45].

## Variables

Keeping this guideline in mind, we selected independent variables for the analysis. To achieve this, we converted RTS, ISS, and time to first CT into continuous variables, while categorical variables were divided into as few levels as possible.

### Study outcome

The study’s outcome was the presence of any opportunity for improvement. An opportunity for improvement presents preventable errors in care with direct adverse effects on patient outcomes or recurrent deviations from safe clinical practice. Examples include technical error, clinical judgement error or any preventable death.

### Patient and process factors

The variables selected for analysis were patient and process factors chosen from the trauma registry, based on the locally used audit filters, previous literature and expert opinion.Categorical factors included sex, mechanical ventilation, on-call hours, ICU length of stay, ASA preinjury, and 30-day mortality. Sex was categorized as male sex or female sex. Mechanical ventilation was stratified into “Not intubated”, “Mechanical ventilation 1-7 days”, “Mechanical ventilation > 7 days” and “Unknown”. On-call hours was defined as arrival to the hospital before 8.00 a.m., after 5 p.m., or during weekends – Saturday or Sunday. ICU length of stay was categorized as “≤ 7 days” or “> 7 days”. ASA preinjury was divided into “ASA 1-2” and “ASA 3-6”. 30-day mortality was determined based on whether the patient was deceased or alive 30 days after injury.

The continuous variables in the analysis were age in years, Revised Trauma Score (RTS), ISS, and time to first computed tomography in minutes. RTS is a measure of severity based on prehospital or emergency department values. In instances where hospital values for systolic blood pressure (SBP), respiratory rate (RR), or Glasgow Come Scale (GCS) scores were missing from the emergency department records, prehospital values were used instead. If the patient was intubated, it was assumed that their GCS was 3 and RR was 0. Subsequently, for the RTS variable, each of the variables GCS, SBP, and RR were converted to RTS codes ranging from 0 to 4 based on their respective values, and the final score was calculated using published coefficients [@71].

## Statistical analysis

We conducted complete case analysis after handling missing values for SBP, RR and GCS as described above. Pearson’s Chi-squared test and the Wilcoxon rank sum test were used to calculate the significance for categorical and continuous variables respectively. Continuous variables are presented using mean values and standard deviations,  and categorical variables are presented using the number of patients within each subgroup and the corresponding subgroup percentages.

Univariate logistic regression was used to determine unadjusted associations and multivariable logistic regression to determine adjusted associations between patient and process factors and opportunities for improvement. Odds ratios (OR) with associated 95% confidence intervals were usecalculated. A p-value of less than 0.05 considered indicative of a significant difference. For continuous variables, ORs were derived by evaluating each unitary step increment. A subgroup analysis of patients who were alive 30 days after the injury was conducted to reduce impact from death as a cause of the association between opportunities for improvement and patient and process factors. All statistical analyses were conducted using R version 4.4 [@19].

# Results

## Participants
Figure 1. Flowchart describing the inclusion and exclusion criteria of the study

```{r, include=FALSE}
source("functions/flowchart.R")
```

```{r, echo=FALSE}
for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert") %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L")
connectGrob(vert$included, excluded1, type = "L")

vert
excluded
excluded1
```

The patient demographics of this study is presented in Figure 1. A total of 14,022 patients were included in the trauma registry and trauma care quality database between 2014 and 2023, out of which 12,278 patients were excluded. Out of all patients, 11,343 were excluded because they were not treated in the ICU, three patients were excluded because they were under 15 years old, six patients were excluded because they were dead on arrival, and 928 patients were excluded because they had not been screened regarding the presence of opportunities for improvement This left us with a total of 1,742 patients eligible for the study. Out of these, 293 patients were excluded because there was missing data in one or several of the patient and process factors, leaving a total of 1,449 patients included in the study.

## Descriptive data

```{r include_scripts, include=FALSE}
source("functions/table1.R")
```

```{r, echo=FALSE}
table2
```

Table 1 present sample characteristics. Most patients were male 1,096 (76%) vs 353 (24%) female patients. There were 846 (58%) patients who were not intubated, 241 (17%) patients who received mechanical ventilation for 1-7 days, and 113 (7,8%) patients who received mechanical ventilation for more than seven days. The number of patients who arrived outside working hours was 1,042 (72%). A total of 1,064 (73%) patients were equal to or less than seven days in the ICU, and 385 (27%) patients were more than seven days in the ICU. A total of 1,080 (75%) patients had an ASA score of 1-2 before arrival to the hospital, and 369 (25%) patients had an ASA score of 3-6 before arrival to the hospital. Amongst our patients, 30-day mortality was at 286 (20%).

When comparing patients with opportunities for improvement to those without, differences emerged across various factors. Specifically, individuals with opportunities for improvement were less likely to require intubation, had higher RTS scores, experienced shorter time to initial CT scans, spent more days in the ICU, and demonstrated a greater likelihood of surviving 30 days post-injury.

## Main results

```{r, include=FALSE}
source("functions/table2.R")
```

```{r, echo=FALSE}
table3_merge
```

Table 2 shows the unadjusted and adjusted associations between patient and process factors and opportunities for improvement. In the unadjusted analysis, the factors that were significantly associated with the presence of opportunities for improvement were: Revised Trauma Score `r inline_text(table3a, variable = RTS)`, > 7 days in the ICU `r inline_text(table3a, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3a, variable = Survival, level = "Alive")`.

In the adjusted analysis the following factors demonstrated significant association with increasing odds of opportunities for improvement: mechanical ventilation > 7 days `r inline_text(table3b, variable = Intubation, level = "Mechanical ventilation > 7 days")`, Revised Trauma Score `r inline_text(table3b, variable = RTS)`, > 7 days in the ICU `r inline_text(table3b, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3b, variable = Survival, level = "Alive")`.

Sex, age, ISS, time to first CT, on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in the unadjusted nor the adjusted analyses.

```{r, include=FALSE}
source("functions/table3.R")
```

```{r, echo=FALSE}
table3alive_merge
```

Table 3 shows the unadjusted and adjusted associations of opportunities for improvement and the selected patient and process factors among the patients who were alive 30 days after injury. In the unadjusted analysis > 7 days in the ICU was significantly associated with opportunities for improvement. In the adjusted analysis increased RTS-score was significantly associated with increased odds of opportunities for improvement.

# Discussion

The key results from this study are that higher RTS, > 7 days length of stay in the ICU and survival 30 days after injury were significantly associated with higher odds of opportunities for improvement. Age, sex, mechanical ventilation, ISS, time to first CT, arrival during on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in either the adjusted or the unadjusted analyses. In the subgroup analysis of patients who were alive 30 days after injury, no patient or process factor was significantly associated with opportunities for improvement in the adjusted analysis.

We found that higher RTS was associated with increased odds of opportunities for improvement. Previous research has shown that patients with relatively better vital parameter values, had a significant association with increased odds of opportunities for improvement in the unadjusted analysis, but not in the adjusted analysis [@12]. These findings could be because patients with lower RTS are often in critical condition upon arrival — upon which emergency interventions are performed immediately. Conversely, patients with higher RTS, indicating relative stability in this patient group upon arrival, may experience delayed interventions.

Similarly, prolonged ICU stay was found to be significantly associated with opportunities for improvement. There are two things to consider regarding this variable. Firstly, patients requiring extended ICU care are often those with complex medical conditions, severe injuries, or critical illness. As a result, their prolonged stay in the ICU may signify a higher baseline risk for experiencing complications, increasing odds of opportunities for improvement. Additionally, prolonged ICU stay increases likelihood of medical errors and adverse events to occur [*REF*]. The ICU environment, while designed to provide advanced medical care, can also be fraught with complexities, including high patient acuity, numerous interventions, and a fast-paced workflow. The longer a patient remains in the ICU, the greater the cumulative exposure to potential errors. Conversely, shorter ICU stays may indicate patient demise before potential mistakes could occur, particularly in cases where mortality was anticipated rather than preventable. This finding is similar to previous research demonstrating prolonged ICU stays as a risk factor for increased morbidity and mortality in trauma patients [@22, @26]. Furthermore, it also corresponds with a retrospective cohort indicating that lower initial illness severity and younger age are associated with earlier discharge, further supporting the link between higher odds of opportunities for improvement and prolonged ICU stays [@25].

The reason for prolonged ICU stays being associated with increased mortality has been investigated by previous studies suggesting that factors such as comorbidities, age, and illness severity upon admission do not fully elucidate the increased long-term mortality observed in this patient cohort. As the duration of ICU stay increases, the primary reason for ICU stay becomes less linked to the patient’s original condition and increasingly tied to persistent chronic illness accompanied by subsequent organ dysfunction. The cause of the organ dysfunction comes from increased exposure to iatrogenic factors and nosocomial infections, which exacerbate pre-existing organ dysfunction or lead to new-onset organ dysfunctions. [@23, @24] Consequently, this might be a potential explanation for the nonsignificance of predictors such as ASA preinjury and ISS in our study. These factors primarily reflect the patient's baseline health status and the initial injury severity, whereas the complications that arise during prolonged ICU stays are more closely associated with the ongoing healthcare interventions and exposures experienced during hospitalization which cannot be measured with ASA preinjury and ISS.

Another factor associated with increased odds of opportunities for improvement was survival 30 days after injury. Previous research has identified an association between increased mortality and the severity of the injury upon admission [@59], suggesting that patients at high risk of 30-day mortality are typically in poor health upon arrival. Consequently, the likelihood of opportunities for improvement decreases as deaths may not be preventable.

We also found that mechanical ventilation more than seven days was significantly associated with decreased odds of opportunities for improvement. Previous research has shown that prolonged mechanical ventilation is associated with increased mortality. This is because mechanical ventilation is associated with severe complications and failure of weaning further increases length of mechanical ventilation as well as ICU length of stay [@65, @66]. Our finding of this association may thus be because of the increased rate of complications and mortality associated with prolonged mechanical ventilation.
*skriv om så det gäller minskad OFI*

Several limitations need to be acknowledged in our study. This was a single-center study and primarily provide insights into the context of the Karolinska University Hospital in Solna. Another limitation is the selection of patients for the morbidity and mortality review, which heavily relied on the local audit filters. The use of audit filters for patient selection for the morbidity and mortality review is associated with high false positive rates, ranging from 24% to 80% [@38]. Even though the use of audit filters is associated with high false positive rates, there is still a risk of misplacement where some patients with opportunities for improvement may have been overlooked. Furthermore, the handling of the variables to the RTS score – namely RR, SBP, and GCS – introduced potential sources of variability. The use of prehospital, emergency department, or intubation-based values for these variables may have affected data accuracy and reproducibility.

# Conclusion
The findings of this study revealed a significant association between higher RTS scores, ICU stay exceeding seven days, mechanical ventilation exceeding seven days, and survival 30 days after injury with increased odds of opportunities for improvement. These factors represent viable targets for new initiatives to further improve the quality of trauma care.

# References