### 2024-06-05 manuscript2.Rmd
5/6

---
title: "Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study"
output: "word_document"
author: "Elin Sun Cao"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based study
### Background
Trauma is a leading cause of mortality and morbidity worldwide. A cornerstone in trauma quality improvement programs is multidisciplinary mortality and morbidity reviews. The purpose of these reviews is to identify opportunities for improvement and corresponding corrective actions. Many patients with severe trauma are admitted to the intensive care unit (ICU), but little is known about opportunities for improvement in this cohort. The aim of this study was to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

### Methods
A registry-based study of trauma patients admitted to the ICU between 2014-2023 at the Karolinska University Hospital in Solna, Sweden, who were reviewed regarding the presence of opportunities for improvement in a mortality and morbidity review. Univariate and multivariable logistic regression was used to determine associations between selected patient and process factors and opportunities for improvement.

### Results
Opportunities for improvement was identified in 143 (9,9%) out of 1449 patients. Mechanical ventilation > 7 days (OR 0.35; 95% CI 0.10, 0.89; p=0.05), Revised Trauma Score (OR 1.21; 95% CI 1.08, 1.37; p=0,002), > 7 days in the ICU (OR 1.93; 95% CI 1.23, 3.03; p=0.004) and survival 30 days after injury (OR 3.29; 95% CI 1.60, 7.54; p=0.002) were significantly associated with opportunities for improvement.

### Conclusion
Several patient and process factors were found to be associated with opportunities for improvement. Notably, patients who were relatively more stable upon arrival and patients with prolonged ICU stays had the highest odds of opportunities for improvement. This underscores the necessity of concentrating more efforts on these specific patient groups.

# Introduction
Trauma — the clinical entity of external injury and the body’s associated response — is a leading cause of mortality and morbidity worldwide [@60][@6]. Patients with major trauma often require intensive care unit (ICU) admission, where managing these patients can be challenging and can result in complications or death [@64][@69]. Although these complications are often expected, the complexity of ICU care necessitates providers to be highly adaptable, which makes the setting prone to mistakes and errors. Peer review is a cornerstone of trauma quality improvement programs aimed at identifying preventable causes of mortality and morbidity in trauma patients [@70], referred to as opportunities for improvement [@42]. The identification of these help derive corrective action plans to improve the quality of trauma care. Previous research on opportunities for improvement in trauma care has often focused on the initial management, identifying areas such as airway and ventilator management, fluid resuscitation, delays to treatment and judgement errors as common opportunities for improvement [@72][@73][@74].

Assessing preventable deaths among ICU-admitted patients poses a significant challenge due to their medical complexity, making it difficult to establish the standard of care for each individual. Moreover, despite receiving optimal treatment, these patients remain at elevated risk for complications and increased mortality rates [@9]. While previous studies have explored various facets of trauma care, including the ICU-setting [@50], there is a notable gap in research regarding error reporting practices in trauma care - especially for trauma patients admitted to the ICU [@64][@12]. Understanding and addressing this gap is essential for improving patient outcomes and optimizing trauma care protocols. Therefore, this study aims to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

# Methods

## Study design

We conducted a retrospective single-center cohort study based on data from the Karolinska University Hospital trauma registry, which is included in the Swedish Trauma registry, and the hospital's local trauma care quality database.

## Setting

The Karolinska University Hospital in Solna is the trauma center designated to receive all severely injured patients in the wider metropolitan area of Stockholm [@17]. The hospital has direct access to radiology, surgery, intensive care, and consultants from relevant specialities [@18]. The care of adult trauma patients is screened for opportunities for improvement using a combination of audit filters and individual review by specialized nurses. Those identified as having a higher likelihood of opportunities for improvement are discussed in a multidisciplinary peer-review held every 6-8 weeks. The presence or absence of opportunities for improvement is decided among the participants of the review and subsequently documented in the trauma care quality database. Before 2017 a non-systematic selection from clinicians involved in the opportunities for improvement process was done by filtering for patients that were thought to have a high risk for opportunities for improvement, but starting in 2017 the care of all patients is screened.

Table 1 illustrates opportunities for improvement divided into categories based on the definitions of opportunity for improvement types during mortality and morbidity reviews at the Karolinska University Hospital in Solna.

```{r, include=FALSE}
#source("functions/ofi_categories.R")
```

## Participants

The trauma registry includes all patients admitted with trauma team activation, regardless of Injury Severity Score, as well as patients admitted without trauma team activation but found to have an Injury Severity Score of more than 9. We included all patients who had been included in the opportunities for improvement screening process between January 1, 2014, and February 1, 2023, and who were also admitted to the ICU. We excluded patients who were younger than 15 years and patients who were dead on arrival. Patients under 15 years of age were not included in the study as their clinical and review pathway differ compared to adults. Patients who were dead on arrival were excluded as opportunities for improvement among these patients is not relevant for the aim of this study.

## Study size

When conducting logistic regression analysis, the number of events per variable equal to 10 is a rule of thumb. If the events per variable falls below this threshold, the results for regression analyses should be interpreted with caution [@45].

## Variables

We categorized and converted the independent variables according to the aforementioned guideline of events per variable to achieve an acceptable multivariate analysis. Specifically, Revised Trauma Score, Injury Severity Score, and time to first computed tomography were treated as continuous variables, while categorical variables were divided into as few levels as possible.

### Patient and process factors

The variables selected for analysis were patient and process factors chosen from the trauma registry, based on the locally used audit filters, previous literature and expert opinion. Categorical factors included sex, mechanical ventilation, on-call hours, ICU length of stay, ASA preinjury, and 30-day mortality. Sex was categorized as male sex or female sex. Mechanical ventilation was stratified into “Not intubated”, “Mechanical ventilation 1-7 days”, “Mechanical ventilation > 7 days” and “Unknown”. On-call hours was defined as arrival to the hospital before 8.00 a.m., after 5 p.m., or during weekends – Saturday or Sunday. ICU length of stay was categorized as “≤ 7 days” or “> 7 days”. American Society of Anesthesiologists (ASA) preinjury was divided into “ASA 1-2” and “ASA 3-6”. 30-day mortality was determined based on whether the patient was deceased or alive 30 days after injury.

The continuous variables in the analysis were age in years, Revised Trauma Score, Injury Severity Score, and time to first computed tomography in minutes. Revised Trauma Score is a measure of severity based on prehospital or emergency department values. In instances where hospital values for systolic blood pressure, respiratory rate, or Glasgow Come Scale scores were missing from the emergency department records, prehospital values were used instead. If the patient was intubated, it was assumed that their Glasgow Coma Scale was three and respiratory rate was zero. Subsequently, for the Revised Trauma Score variable, each of the variables Glasgow Coma Scale, systolic blood pressure, and respiratory rate were converted to Revised Trauma Score codes ranging from 0 to 4 based on their respective values, and the final score was calculated using published coefficients [@71].

### Study outcome

The study’s outcome was the presence of any opportunity for improvement. An opportunity for improvement presents preventable errors in care with direct adverse effects on patient outcomes or recurrent deviations from safe clinical practice. Examples include technical error, clinical judgement error or any preventable death.

## Statistical analysis

We conducted complete case analysis after handling missing values for systolic blood pressure, respiratory rate and Glasgow Coma Scale as described above. Pearson’s Chi-squared test and the Wilcoxon rank sum test were used to calculate the significance for categorical and continuous variables respectively. Continuous variables are presented using mean values and standard deviations, and categorical variables are presented with the number of patients in each subgroup and their respective percentages.

Univariate logistic regression was used to determine unadjusted associations and multivariable logistic regression to determine adjusted associations between patient and process factors and opportunities for improvement. Odds ratios (OR) with associated 95% confidence intervals were calculated. A p-value of less than 0.05 considered indicative of a significant difference. For continuous variables, ORs were derived by evaluating each unitary step increment. A subgroup analysis of patients who were alive and dead 30 days after the injury was conducted to reduce impact from death as a cause of the association between opportunities for improvement and patient and process factors. All statistical analyses were conducted using R version 4.4 [@19].

# Results

## Participants
Figure 1. Flowchart describing the inclusion and exclusion criteria of the study

```{r, include=FALSE}
source("functions/flowchart.R")
```

```{r, echo=FALSE}
for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert", arrow = small_arrow) %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L", arrow = small_arrow)
connectGrob(vert$included, excluded1, type = "L", arrow = small_arrow)

vert
excluded
excluded1
```

The patient demographics of this study is presented in Figure 1. A total of 14,022 patients were included in the trauma registry and trauma care quality database between 2014 and 2023, out of which 12,278 patients were excluded. Out of all patients, 11,343 were excluded because they were not treated in the ICU, three patients were excluded because they were under 15 years old, six patients were excluded because they were dead on arrival, and 928 patients were excluded because they had not been screened regarding the presence of opportunities for improvement This left us with a total of 1,742 patients eligible for the study. Out of these, 293 patients were excluded because there was missing data in one or several of the patient and process factors, leaving a total of 1,449 patients included in the study. The variable with the highest number of missing values was Revised Trauma Score, with 207 instances of missing data.

## Descriptive data

```{r include_scripts, include=FALSE}
source("functions/table1.R")
```

```{r, echo=FALSE}
table2
```

Table 2 present sample characteristics. Most patients were male 1,096 (76%) vs 353 (24%). There were 846 (58%) patients who were not intubated, 241 (17%) patients who received mechanical ventilation for 1-7 days, and 113 (7,8%) patients who received mechanical ventilation for more than 7 days. The number of patients arriving outside working hours was 1,042 (72%). A total of 1,064 (73%) patients were equal to or less than 7 days in the ICU vs 385 (27%) patients whose ICU-stay exceeded 7 days. A total of 1,080 (75%) patients had an ASA score of 1-2 before arrival to the hospital, and 369 (25%) patients had an ASA score of 3-6 before arrival to the hospital. Amongst our patients, 30-day mortality was at 286 (20%).

When comparing patients with opportunities for improvement to those without, differences emerged across various factors. Specifically, individuals with opportunities for improvement were less likely to require intubation, had higher Revised Trauma Score, experienced shorter time to initial computed tomography scans, spent more days in the ICU, and demonstrated a greater likelihood of surviving 30 days post-injury.

## Main results

```{r, include=FALSE}
source("functions/table2.R")
```

```{r, echo=FALSE}
table3_merge
```

Table 3 shows the unadjusted and adjusted associations between patient and process factors and opportunities for improvement. In the unadjusted analysis, the factors that were significantly associated with the presence of opportunities for improvement were: Revised Trauma Score `r inline_text(table3a, variable = RTS)`, > 7 days in the ICU `r inline_text(table3a, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3a, variable = Survival, level = "Alive")`.

In the adjusted analysis the following factors demonstrated significant association with opportunities for improvement: mechanical ventilation > 7 days `r inline_text(table3b, variable = Intubation, level = "Mechanical ventilation > 7 days")`, Revised Trauma Score `r inline_text(table3b, variable = RTS)`, > 7 days in the ICU `r inline_text(table3b, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3b, variable = Survival, level = "Alive")`.

Sex, age, Injury Severity Score, time to first computed tomography, on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in the unadjusted nor the adjusted analyses.

```{r, include=FALSE}
source("functions/table3.R")
```

```{r, echo=FALSE}
table3b_merge
```

Table 4 shows the adjusted associations of opportunities for improvement and the selected patient and process factors among the patients who were alive and dead respectively 30 days after injury. Revised Trauma Score was significantly associated with increased odds of OFI for both patients who were alive `r inline_text(table3aalive, variable = RTS)` and deceased `r inline_text(table3bdead, variable = RTS)` 30 days after injury. More than 7 days in the ICU was significantly associated with increased odds of opportunities for improvement among patients who were deceased 30 days after injury `r inline_text(table3bdead, variable = daysinICU, level = "> 7 days")`.

# Discussion

The key results from this study are that higher Revised Trauma Score, > 7 days length of stay in the ICU, and survival 30 days after injury were significantly associated with higher odds of opportunities for improvement. More than 7 days on mechanical ventilation was associated with decreased odds of opportunities for improvement. Age, sex, Injury Severity Score, time to first computed tomography, arrival during on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in neither the adjusted nor the unadjusted analyses. In the analysis of patients who were alive and deceased 30 days after injury increased Revised Trauma Score was significantly associated with higher odds of opportunities for improvement in both patient groups. Among the patients who deceased 30 days after injury, more than 7 days in the ICU was significantly associated with higher odds of opportunities for improvement.

Our study found a significant association between higher Revised Trauma Score and higher odds of opportunities for improvement, indicating that among very severely injured patients that are admitted to the ICU, patients in a less severe condition were at higher odds of opportunities for improvement. Patients with lower Revised Trauma Score are often in critical condition upon arrival — upon which emergency interventions are performed immediately. In contrast, patients with higher Revised Trauma Score, indicating relative stability upon arrival, may experience delays in interventions overall and might receive less rigorous monitoring. This aligns with findings from previous research conducted on trauma patients from the Karolinska University Hospital Solna, where it was noted that patients with more stable vital parameters were more likely to have increased odds of opportunities for improvement [@12].

Prolonged ICU stay was also found to be significantly associated with opportunities for improvement. This variable warrants careful consideration for two primary reasons. Firstly, patients requiring extended ICU care are often those with complex medical conditions or severe injuries. As a result, their prolonged ICU stay may signify a higher baseline risk for experiencing complications, increasing odds of opportunities for improvement. The ICU environment, while designed to provide advanced medical care, can also be fraught with complexities, including high patient acuity, numerous interventions, and a fast-paced workflow, all contributing to a higher cumulative exposure to potential errors the longer a patient remains in the ICU. Additionally, complications or opportunities for improvement might themselves lead to longer ICU stays. Conversely, shorter ICU stays may indicate patient demise before potential mistakes could occur, particularly in cases where mortality was anticipated rather than preventable. This finding is similar to previous research demonstrating prolonged ICU stays as a risk factor for increased mortality and morbidity in trauma patients [@22, @26]. Furthermore, it also corresponds with a retrospective cohort indicating that lower initial illness severity and younger age are associated with earlier discharge, further supporting the link between higher odds of opportunities for improvement and prolonged ICU stays [@25].

The reason for prolonged ICU stays being associated with increased mortality has been investigated by previous studies suggesting that factors such as comorbidities, age, and illness severity upon admission do not fully elucidate the increased long-term mortality observed in this patient cohort. As the duration of ICU stay increases, the primary reason for ICU stay becomes less linked to the patient’s original condition and increasingly tied to persistent chronic illness accompanied by subsequent organ dysfunction. The cause of the organ dysfunction comes from increased exposure to iatrogenic factors and nosocomial infections, which exacerbate pre-existing organ dysfunction or lead to new-onset organ dysfunctions. [@23, @24] Consequently, this might be a potential explanation for the nonsignificance of predictors such as ASA preinjury and Injury Severity Score in our study. These factors primarily reflect the patient's baseline health status and the initial injury severity, whereas the complications that arise during prolonged ICU stays are more closely associated with the ongoing healthcare interventions and exposures experienced during hospitalization which cannot be measured with ASA preinjury and Injury Severity Score.

Another factor associated with increased odds of opportunities for improvement was survival 30 days after injury. Previous research has identified an association between increased mortality and the severity of the injury upon admission [@59], suggesting that patients at high risk of 30-day mortality are typically in poor health upon arrival. Consequently, the likelihood of opportunities for improvement decreases as deaths may not be preventable.

We found that mechanical ventilation more than 7 days was significantly associated with decreased odds of opportunities for improvement. Previous research has shown that prolonged mechanical ventilation is associated with increased mortality. This is because mechanical ventilation is associated with severe complications and failure of weaning further increases length of mechanical ventilation as well as ICU length of stay [@65][@66]. Our finding of decreased odds of OFI may be attributed to the fact that, despite the increased rate of complications and mortality associated with prolonged mechanical ventilation, it was the appropriate intervention for these patients.

One of the strengths of this study is its reliance on registry-based data, ensuring a high level of data quality and reliability. The use of pre-defined variables standardized their definitions, enhancing the potential for similar studies to replicate these findings. Additionally, the study’s specific inclusion criteria facilitated the formation of a well-defined cohort, further contributing to the reliability and reproducibility of the study.

However, several limitations need to be acknowledged in our study. This was a single-center study and primarily provide insights into the context of the Karolinska University Hospital in Solna. Another limitation is the selection of patients for the mortality and mortbidity review, which heavily relied on the local audit filters. The use of audit filters for patient selection for the mortality and morbidity review is associated with high false positive rates, ranging from 24% to 80% [@38]. Even though the use of audit filters is associated with high false positive rates, there is still a risk of misplacement where some patients with opportunities for improvement may have been overlooked. Furthermore, the handling of the variables to the Revised Trauma Score – namely respiratory rate, systolic blood pressure, and Glasgow Coma Scale – introduced potential sources of variability. The use of prehospital, emergency department, or intubation-based values for these variables may have affected data accuracy and reproducibility.

# Conclusion

The findings revealed a significant association between increased odds of opportunities for improvement and higher Revised Trauma Score, ICU length of stay exceeding 7 days, and survival 30 days after injury, as well as a significant association between decreased odds of opportunities for improvement and more than 7 days on mechanical ventilation. The results highlight the importance of prioritizing these groups of patients for closer monitoring and implementing targeted interventions to prevent errors in care. Further research on opportunities for improvement among ICU-admitted patients is needed to expand our understanding of this aspect of trauma care. The primary limitation of this study was its restricted sample size, highlighting the need for larger-scale investigations. Conducting multi-center studies encompassing extensive databases or merging multiple datasets could provide more comprehensive insights into opportunities for improvement and contribute to enhancing Trauma Quality Improvement Programs on a broader scale. However, it is important to acknowledge that multi-center studies may compromise internal validity, reducing their applicability to specific centers.

Building upon the findings of this study, future research should focus on specific variables such as ICU length of stay and the individual assessment of vital parameters like Glasgow Coma Scale, respiratory rate, and systolic blood pressure, rather than relying solely on composite scores like Revised Trauma Score. Analyzing these vital signs separately may reveal which parameter holds greater predictive value for opportunities for improvement. Also, qualitative studies could provide an in-depth understanding of why physiologically more stable patients are more likely to have opportunities for improvement.

Given the significant association between higher Revised Trauma Score and increased odds of opportunities for improvement, attention should be directed towards trauma patients who appear stable upon admission but may deteriorate over time. Future research should aim to identify early indicators of worsening conditions in these patients, facilitating more targeted resource allocation and intervention strategies.

In contrast, Injury Severity Score did not demonstrate a correlation with opportunities for improvement in this study. Subsequent studies could explore alternative scoring systems such as New Injury Severity Score, Revised Trauma Score, and ASA scores, or combinations of physiological and anatomical scoring systems to determine the most effective predictor of opportunities for improvement in trauma patients.

Overall, future research endeavors should strive to address these gaps in knowledge to optimize trauma care outcomes and contribute to the ongoing improvement of trauma quality improvement initiatives.

# References