### 2024-11-23 manuscript2.Rmd
23/11

---
title: "Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based cohort study"
output: "html_document"
author: "Elin Sun Cao"
bibliography: bibliography.bib
csl: vancouver.csl
editor_options: 
  markdown: 
    wrap: 72
---

## Opportunities for improvement in adult trauma patients admitted to the intensive care unit: A registry-based cohort study
## Abstract
### Background
Trauma is a leading cause of mortality and morbidity worldwide. A cornerstone in trauma quality improvement programs is multidisciplinary mortality and morbidity reviews. The purpose of these reviews is to identify opportunities for improvement and corresponding corrective actions. Many patients with severe trauma are admitted to the intensive care unit (ICU), but little is known about opportunities for improvement in this cohort. The aim of this study was to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

### Methods
A registry-based study of trauma patients admitted to the ICU between 2014-2023 at the Karolinska University Hospital in Solna, Sweden, who were reviewed regarding the presence of opportunities for improvement in a mortality and morbidity review. Univariate and multivariable logistic regression was used to determine associations between selected patient and process factors and opportunities for improvement.

### Results
Opportunities for improvement was identified in 143 (9,9%) out of 1449 patients.Revised Trauma Score , > 7 days in the ICU  and survival 30 days after injury  were significantly associated with increased opportunities for improvement.

### Conclusion
The care of patients who are physiologically more stable on admission to the ICU is more likely to be associated with opportunities for improvement. This association is particularly pronounced in patients who ultimately die. Additionally, higher odds of opportunities for improvement in survivors and those with prolonged ICU stays suggest that extended care in complex cases may both heighten error risk and lead to longer hospitalisations.

# Introduction
Trauma — the clinical entity of external injury and the body’s associated response — is a leading cause of mortality and morbidity worldwide [@60; @6]. Patients with major trauma often require intensive care unit (ICU) admission, where managing these patients can be challenging and result in complications or death [@64; @69]. Although these complications are often expected, the complexity of ICU care makes the setting prone to mistakes and errors.

Peer review is a cornerstone of trauma quality improvement programs aimed at identifying preventable causes of mortality and morbidity in trauma patients [@70], referred to as opportunities for improvement [@42]. Previous research on opportunities for improvement was recently classified in a systematic review that identified areas for improvement such as inadequate monitoring, team communication, assessment of injury and procedure-related issues [@75].

In contrast, little is known about opportunities for improvement in the care of trauma patients in the ICU [@64; @12]. Assessing opportunities for improvement in these patients is challenging because it is difficult to establish the standard of care for each individual and despite optimal treatment, these patients are at risk for complications and death [@9].  This study aims to characterize opportunities for improvement in adult trauma patients admitted to the ICU and assess how patient and process factors are associated with opportunities for improvement in these patients.

# Methods

## Study design and data sources

We conducted a retrospective single-center cohort study based on data from the Karolinska University Hospital trauma registry, and the hospital’s local trauma care quality database. The Karolinska University Hospital trauma registry is part of the Swedish national trauma registry SweTrau which follows the Utstein template [@82]. The registry includes all patients admitted with trauma team activation, regardless of Injury Severity Score, as well as patients admitted without trauma team activation but found to have an Injury Severity Score of more than 9 [@76]. The trauma care quality database includes data collected as part of the local peer review process. Screening for opportunities for improvement in selected patient cases was initiated in 2013. Since 2017 all patients have been included in the screening process.

## Setting

The Karolinska University Hospital in Solna is the trauma center designated to receive all severely injured patients in the wider metropolitan area of Stockholm [@17]. The hospital has direct access to radiology, surgery, intensive care, and consultants from relevant specialities [@18]. The care of adult trauma patients is screened for opportunities for improvement using a combination of audit filters and individual review by specialized nurses. All deceased patients, as well as those deemed as having a higher probability of opportunities for improvement, are reviewed in a multidisciplinary peer-review conducted every 6–8 weeks. The presence or absence of opportunities for improvement, including any corrective actions, is reached through consensus among the peer-review participants and subsequently recorded in the trauma care quality database. Decisions regarding the presence of preventable deaths are also made during the peer-review process.. The different ICU:s at the Karolinska University Hospital are all Level 3 as defined by the Intensive Care Society [@83].

## Participants

We included all patients who had been included in the opportunities for improvement screening process between January 1, 2013, and February 28, 2023, and who were also admitted to the ICU. We excluded patients who were younger than 15 years and patients who were dead on arrival. Patients under 15 years of age were not included in the study as their clinical and review pathway differ compared to adults. Patients who were dead on arrival were excluded as opportunities for improvement among these patients is not relevant for the aim of this study.

## Variables

### Outcome

The outcome was the presence of any opportunity for improvement during the hospital stay, as decided by consensus during a multidisciplinary peer-review meeting. An opportunity for improvement represents preventable events in the care of patients with adverse effects on their outcomes, or recurrent deviations from safe clinical practice, which are classified into the categories missed diagnosis, delay in treatment, clinical judgment errors, inadequate protocols, inadequate resources, other errors, and preventable deaths.

### Patient and process factors

The variables selected for analysis were patient and process factors chosen from the trauma registry, based on the locally used audit filters, previous literature and expert opinion. The categorical factors included sex [@30; @31], mechanical ventilation [@66; @78], on-call hours [@79], ICU length of stay [@22; @26], American Society of Anesthesiologists preinjury score [@33], and 30-day mortality. Sex was categorized as male sex or female sex. Mechanical ventilation was stratified into “Not intubated”, “Mechanical ventilation 1-7 days”, “Mechanical ventilation > 7 days” and “Unknown”. On-call hours was defined as arrival to the hospital before 8.00 a.m., after 5 p.m., or during weekends – Saturday or Sunday. ICU length of stay was categorized as “≤ 7 days” or “> 7 days”. American Society of Anesthesiologists preinjury score was divided into “ASA 1-2” and “ASA 3-5”. We categorized variables to limit the number of independent variables in the multivariable analysis to adhere to the rule of thumb of ten patients with the outcome per variable [@45].

The continuous variables in the analysis were age in years [@81], Revised Trauma Score [@11], Injury Severity Score [@77], and time to first computed tomography in minutes [@80]. The Revised Trauma Score is a measure of severity based on prehospital or emergency department values. In instances where hospital values for systolic blood pressure, respiratory rate, or Glasgow Coma Scale scores were missing from the emergency department records, prehospital values were used instead. If it was known that the patient was intubated, it was assumed that their Glasgow Coma Scale was three and respiratory rate was zero when calculating the Revised Trauma Score variable. Subsequently, each of the variables Glasgow Coma Scale, systolic blood pressure, and respiratory rate were converted to Revised Trauma Score codes ranging from 0 to 4 based on their respective values, and the final score was calculated using published coefficients [@71].

## Statistical analysis

We conducted a complete case analysis after handling missing values for systolic blood pressure, respiratory rate and Glasgow Coma Scale as described above. Continuous variables are presented using mean values and standard deviations, and categorical variables are presented with the number of patients in each subgroup and their respective percentages.

We used logistic regression to determine unadjusted and adjusted associations between patient and process factors and opportunities for improvement. Odds ratios (OR) with associated 95% confidence intervals were calculated. A p-value of less than 0.05 were considered indicative of a significant difference. We considered 30-day mortality as a potential effect modifier and therefore conducted a stratified analysis in which the associations between patient and process factors and opportunities for improvement were estimated separately in patients who survived and patients who died within 30 days. All statistical analyses were conducted using R version 4.4 [@19].

# Results

## Participants
Figure 1. Flowchart describing the inclusion and exclusion criteria of the study

```{r, include=FALSE}
source("functions/flowchart.R")
```

```{r, echo=FALSE}
for (i in 1:(length(vert) - 1)) {
  connectGrob(vert[[i]], vert[[i + 1]], type = "vert", arrow = small_arrow) %>%
    print
}

connectGrob(vert$eligible, excluded, type = "L", arrow = small_arrow)
connectGrob(vert$included, excluded1, type = "L", arrow = small_arrow)

vert
excluded
excluded1
```

A total of 14,022 patients were included in the trauma registry and trauma care quality database between 2014 and 2023, out of which 12,278 patients were excluded, as shown in Figure 1. Out of all patients, 11,343 were excluded because they were not treated in the ICU, three patients were excluded because they were under 15 years old, six patients were excluded because they were dead on arrival, and 928 patients were excluded because they had not been screened regarding the presence of opportunities for improvement. This left us with a total of 1,742 patients eligible for the study. Out of these, 293 patients were excluded because there was missing data in one or several of the patient and process factors, leaving a total of 1,449 patients included in the study. The variable with the highest number of missing values was Revised Trauma Score, with 207 instances of missing data.

## Descriptive data

```{r include_scripts, include=FALSE}
source("functions/table1.R")
```

```{r, echo=FALSE}
table2
```

Table 1 presents sample characteristics. Most patients were male 1,096 (76%) vs 353 (24%). Of the 354 (42%) patients who were intubated, 241 (17%) patients received mechanical ventilation for 1-7 days, while 113 (7,8%) patients received mechanical ventilation for more than 7 days. The number of patients arriving outside working hours was 1,042 (72%). A total of 1,064 (73%) patients were equal to or less than 7 days in the ICU vs 385 (27%) patients whose ICU-stay exceeded 7 days. A total of 1,080 (75%) patients had an ASA score of 1-2 before arrival to the hospital, and 369 (25%) patients had an ASA score of 3-6 before arrival to the hospital. Amongst our patients, 30-day mortality was at 286 (20%). Of these, 92 patients died within 24 hours, and a total of 144 patients died within 48 hours of hospital arrival. When comparing patients with opportunities for improvement to those without, patients with opportunities for improvement were less likely to require intubation, had higher Revised Trauma Score, experienced shorter time to initial computed tomography scans, spent more days in the ICU, and demonstrated a greater likelihood of surviving 30 days post-injury.

```{r, include=FALSE}
source("functions/ofi_categories.R")
```
```{r, echo=FALSE}
table_ofic
```

Table 2 illustrates opportunities for improvement divided into categories based on the definitions of opportunity for improvement types during mortality and morbidity reviews at the Karolinska University Hospital in Solna. The most common category of opportunity for improvement in the patient cohort of this study was clinical judgement error, followed by delay in treatment.

## Main results

```{r, include=FALSE}
source("functions/table2.R")
```

```{r, echo=FALSE}
table3_merge
```

Table 3 shows the unadjusted and adjusted associations between patient and process factors and opportunities for improvement. In the unadjusted analysis, Revised Trauma Score `r inline_text(table3a, variable = RTS)`, > 7 days in the ICU `r inline_text(table3a, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3a, variable = Survival, level = "Alive")`, were all significantly associated with opportunities for improvement. These variables remained significantly associated with opportunities for improvement in the adjusted analysis, with similar effect magnitudes: Revised Trauma Score `r inline_text(table3b, variable = RTS)`, > 7 days in the ICU `r inline_text(table3b, variable = daysinICU, level = "> 7 days")` and survival 30 days after injury `r inline_text(table3b, variable = Survival, level = "Alive")`. Sex, age, mechanical ventilation, Injury Severity Score, time to first computed tomography, on-call hours, and ASA preinjury were not significantly associated with opportunities for improvement in the unadjusted nor the adjusted analyses.

```{r, include=FALSE}
source("functions/table3.R")
```

```{r, echo=FALSE}
table3b_merge
```

Table 4 shows the results of the stratified analysis with adjusted associations between patient and process factors and opportunities for improvement among the patients who were alive and dead 30 days after injury respectively. The Revised Trauma Score was significantly associated with increased odds of OFI for both patients who were alive `r inline_text(table3aalive, variable = RTS)` and dead `r inline_text(table3bdead, variable = RTS)` 30 days after injury respectively. Time to first computed tomography was significantly associated with opportunities for improvement in patients who were deceased 30 days after injury `r inline_text(table3bdead, variable = TimeFCT)`.

# Discussion

The key results from this study are that higher Revised Trauma Score, > 7 days length of stay in the ICU, and survival 30 days after injury were significantly associated with higher odds of opportunities for improvement. In the stratified analysis based on survival until 30 days after injury, only the Revised Trauma Score was significantly associated with higher odds of opportunities for improvement in both patient groups. Among the patients who were dead 30 days after injury, > 7 days in the ICU was significantly associated with higher odds of opportunities for improvement in addition to RTS. While these findings cannot be used to proactively identify patients at risk of medical errors or adverse events, they indicate that patients who are physiologically more stable on arrival, particularly those who ultimately die, and those who stay more than one week in the ICU are potential target groups for a focused review process, aiming to identify interventions to improve the quality of their care.

The positive association between the Revised Trauma Score and opportunities for improvement, means that among the very severely injured patients admitted to the ICU, patients in a less severe condition on admission were at higher odds of opportunities for improvement. One hypothetical explanation is that patients with higher Revised Trauma Score, indicating relative stability upon arrival, may experience delays in interventions overall. This aligns with findings from previous research conducted on all trauma patients from the Karolinska University Hospital Solna, where it was noted that patients with more stable vital parameters were more likely to have opportunities for improvement [@12].

Prolonged ICU stay was also found to be significantly associated with opportunities for improvement. This variable warrants careful consideration for two primary reasons. Firstly, patients requiring extended ICU care are often those with complex medical conditions or severe injuries. As a result, their prolonged ICU stay may signify a higher baseline risk for experiencing complications, increasing odds of opportunities for improvement. The ICU environment, while designed to provide advanced medical care, can also be fraught with complexities, including high patient acuity, numerous interventions, and a fast-paced workflow, all contributing to a higher cumulative exposure to potential errors the longer a patient remains in the ICU.

Additionally, complications or opportunities for improvement might themselves lead to longer ICU stays. Conversely, shorter ICU stays may indicate patient demise before potential mistakes could occur, particularly in cases where mortality was anticipated. This is one potential explanation for why death was negatively associated with opportunities for improvement in our study. This finding is similar to previous research demonstrating prolonged ICU stays as a risk factor for increased mortality and morbidity in trauma patients [@22, @26].  Furthermore, it also corresponds with a retrospective cohort indicating that lower initial illness severity and younger age are associated with earlier discharge, further supporting the link between higher odds of opportunities for improvement and prolonged ICU stays [@25].

The reason for prolonged ICU stays being associated with increased mortality has been investigated by previous studies suggesting that factors such as comorbidities, age, and illness severity upon admission do not fully explain the increased long-term mortality observed in this patient cohort. As the duration of ICU stay increases, the primary reason for ICU stay becomes less linked to the patient’s original condition and increasingly tied to persistent chronic illness accompanied by subsequent organ dysfunction. The cause of the organ dysfunction comes from increased exposure to iatrogenic factors and nosocomial infections, which exacerbate pre-existing organ dysfunction or lead to new-onset organ dysfunctions [@23, @24]. Consequently, this might be a potential explanation for the nonsignificance of predictors such as ASA preinjury and Injury Severity Score in our study. These factors primarily reflect the patient’s baseline health status and the initial injury severity, whereas the complications that arise during prolonged ICU stays are more closely associated with the ongoing healthcare interventions and exposures experienced during hospitalization which cannot be measured with ASA preinjury and Injury Severity Score.

One of the strengths of this study is its reliance on registry-based data, ensuring a high level of data quality and reliability. The use of pre-defined variables standardized their definitions, enhancing the potential for similar studies to replicate these findings. Additionally, the study’s specific inclusion criteria facilitated the formation of a well-defined cohort, further contributing to the reliability and reproducibility of the study.

However, several limitations need to be acknowledged in our study. This was a single-center study and primarily provide insights into the context of the Karolinska University Hospital in Solna. Another limitation is the selection of patients for the mortality and mortbidity review, which heavily relied on the local audit filters. The use of audit filters for patient selection for the mortality and morbidity review is associated with high false positive rates, ranging from 24% to 80% [@38]. Even though the use of audit filters is associated with high false positive rates, there is still a risk of misplacement where some patients with opportunities for improvement may have been overlooked. Furthermore, the handling of the variables to the Revised Trauma Score – namely respiratory rate, systolic blood pressure, and Glasgow Coma Scale – introduced potential sources of variability. The use of prehospital, emergency department, or intubation-based values for these variables may have affected data accuracy and reproducibility. Finally, the outcome opportunities for improvement is a consensus decision made in the local trauma quality improvement process. As such, these decisions have not been externally reviewed.

# Conclusion

The findings revealed a significant association between increased odds of opportunities for improvement and higher Revised Trauma Score, ICU length of stay longer than 7 days, and survival 30 days after injury. These findings indicate that patients who are physiologically more stable on arrival but stay more than one week in the ICU are a potential target group for a focused review process, aiming to identify interventions to improve the quality of their care.

# References